plackett_luce_monster

### 2022-07-07 inst/plackett_luce_monster.Rmd
more work on red bus/blue bus
cheering and an application to red bus-blue bus

* Runner $i$ might hate runner $j$.  If $i$ finishes, he might expend
  his spare energy hindering $j$ (by throwing stones or booing , for
  example).  Note that equivalence classes of mutual hatred might
  exist.

@@

@@

 231: $\frac{p_2}{p_2+p_3+p_1}\cdot\frac{p_3}{p_3+p_1+M}$

 312: $\frac{p_3}{p_3+p_1+p_2}\cdot\frac{p_1}{p_2+p_1}$

 321: $\frac{p_3}{p_3+p_2+p_1}\cdot\frac{p_2}{p_2+p_1}$

@@

>>>>

* Runner $i$ might hate runner $j$.  If $i$ finishes, he might expend
  his spare energy hindering $j$ (by throwing stones or booing, for
  example).  Note that equivalence classes of mutual hatred might
  exist.

@@

The runner metaphor is not the only one.  We can apply the resulting
likelihood functions to other cases including the red bus-blue bus
problem (see `inst/red_bus_blue_bus.Rmd` for more discussion).

@@

 231: $\frac{p_2}{p_2+p_3+p_1}\cdot\frac{p_3}{p_3+p_1+M}\qquad\frac{p_2}{p_2+p_1+p_3}\cdot\frac{\lambda p_1}{p_3+\lambda p_1}$

 312: $\frac{p_3}{p_3+p_1+p_2}\cdot\frac{p_1}{p_2+p_1}\qquad\frac{p_3}{p_3+p_1+p_2}\cdot\frac{\lambda p_1}{p_3+p_1}$

 321: $\frac{p_3}{p_3+p_2+p_1}\cdot\frac{p_2}{p_2+p_1}\qquad\frac{p_3}{p_3+p_2+p_1}\cdot\frac{p_2}{p_2+p_1}$

@@

# Package idiom

We can investigate red bus-blue bus phenomenon (as discussed, in a
slightly different context, in `inst/red_bus_blue_bus.Rmd`).  Here, we
consider a person who is given the choice of five transport methods:

* `C`, car
* `T`, train
* `RB` a red bus
* `BB` a blue bus
* `W` walking

Now, he does not really care what colour the bus is.  If we ask him to
rank his options, it is highly probable that he will put `RB` and `BB`
consecutively (because they are essentially indistinguishable).  Can
we quantify the strength of this effect?  To do this, we define a
bespoke function `RB_BB_LF()` which returns a `hyper3` log-likelihood
function corresponding to repeated observations of our commuter's
reported ranks for the five options:

```{r,defineredbuslf}
`RB_BB_LF` <- function(s){
    ec <- c(C=1,T=2,RB=3,BB=3,W=4) # equivalence classes
    h <- c(1,1,s,1)                # strength of support
    (
        cheering3(v=c("RB","BB","C" ,"T","W"),e=ec,h=h)*7 + 
        cheering3(v=c("BB","RB","T" ,"C","W"),e=ec,h=h)*4 + 
        cheering3(v=c("T" ,"BB","RB","C","W"),e=ec,h=h)*5 + 
        cheering3(v=c("W" ,"BB","RB","T","C"),e=ec,h=h)*8 + 
        cheering3(v=c("C" ,"RB","BB","W","T"),e=ec,h=h)*9 + 
        cheering3(v=c("BB","C" ,"RB","T","W"),e=ec,h=h)*5
    )
}
```

Above, we see from the function body that he reported `RB,BB,C,T,W`
seven times, `BB,RB,T,C,W` four times, and so on; perhaps his ranking
depends on the weather or how tired he is on any given day.  Observe
that in almost every case he ranks `RB` and `BB` consecutively.
Function `RB_BB_LF()` takes argument `s` that quantifies the perceived
similarity between `RB` and `BB`.  For example:

```{r,examplerbbblf,cache=TRUE}
(H <- RB_BB_LF(s=1.8888))
(mH <- maxp(H,n=1))
```

Now to find a profile likelihood function for `s`:

```{r,label=proflikeredbus,cache=TRUE}
o <- function(s){maxp(RB_BB_LF(s),give=TRUE,n=1)$likes} # optimand
s <- exp(seq(from=log(1.3),to=log(37),len=17)) # putative similarity measures
L <- sapply(s,o)
L <- L-max(L)
```

We can plot these:

```{r,plotproflike}
plot(s,L,type="b")
abline(h=c(0,-2))
abline(v=1)
plot(log(s),L,type="b")
abline(h=c(0,-2))
abline(v=0)
```

And formally maximize the likelihood:

```{r,label=maxlikesim,cache=TRUE}
(osup <- optimize(o,c(6,10),maximum=TRUE))
```

So a likelihood ratio test of the null that $S=1$ would be:

```{r,label=LRT_RB,cache=TRUE}
(suppdif <- o(osup$maximum) - o(1))
```

Falling just shy of Edwards's two-units-of-support criterion.
